### 2024-04-22 bibliography.tex
added new bibtexs and working principle finished

\ifboolexpr{bool{jacowbiblatex}}%
{\printbibliography}{

    \begin{thebibliography}{99}

        % http://cds.cern.ch/record/782076/files/CERN-2004-003-V1.pdf?version=2
        \bibitem{LHC_report} BR\"UNING, O. \emph{LHC design report Volume I The LHC
            Main Ring}. Geneva: European Organization for Nuclear Research, c2004-. CERN
            (Series), 2004-003. ISBN 92-9083-224-0

        \bibitem{beam_dump_system} Filippini, R., Carlier, E., Ducimetière, L., Goddard,
            B. \& Uythoven, J. \emph{Reliability Analysis of the LHC Beam Dumping System}.
            (2005,6), \url{https://cds.cern.ch/record/841092}

        \bibitem{particles_in_ag}Meddahi, M., Bart Pedersen, S., Boccardi, A.,
            Butterworth, A., Fisher, A., Gianfelice-Wendt, E., Goddard, B., Hemelsoet,
            G., Höfle, W., Jacquet, D., Jaussi, M., Kain, V., Lefevre, T., Shaposhnikova,
            E., Uythoven, J. \& Valuch, D. LHC Abort Gap Monitoring and Cleaning.
            (2010,6), \url{https://cds.cern.ch/record/1277630}

        \bibitem{abort_gap_monitor}De Santis, S., Beche, J., Byrd, J., Placidi, M.,
            Turner, W. \& Zolotorev, M. \emph{Development of an Abort Gap Monitor for the
            Large Hadron Collider}.  (2004), \url{https://cds.cern.ch/record/823343}

        \bibitem{ttc_distribution} Taylor, B. \emph{TTC distribution for LHC detectors}.
            {\em IEEE Trans. Nucl. Sci.}. \textbf{45} pp. 821-828 (1998),
            \url{https://cds.cern.ch/record/363087}

        \bibitem{bsra_operation}Bart Pedersen, S., Lefevre, T., Bravin, E., Boccardi, A.,
          Goldblatt, A., Jeff, A., Roncarolo, F. \& Fisher, A. \emph{First Operation of
          the Abort Gap Monitors for LHC}.  (2010,5), \url{http://cds.cern.ch/record/1272172}

        \bibitem{high_sensitivity_measurement} C. Fischer, \emph{High Sensitivity
            Measurement of the Beam Longitudinal Distribution of the LHC Beams}, 
            (2003,1), LHC-BP-ES-0005, rev. 2.0,
            \url{https://edms.cern.ch/document/328145/2.0}

        \bibitem{Fisher_bsra}Fisher, A. \emph{Expected Performance of the LHC
            Synchrotron-Light Telescope (BSRT) and Abort-Gap Monitor (BSRA).}
            (CERN,2010), \url{https://cds.cern.ch/record/1261191}

        % https://cds.cern.ch/record/2295107/files/1638883_171-176.pdf
        \bibitem{machine_protect}Bravin, E., Belohrad, D., Calvo, E. \& Mazzoni, S. Beam
            instrumentation for machine protection.  (2014), \url{https://cds.cern.ch/record/2295107}

        % https://cds.cern.ch/record/1261191/files/performance-note-014.pdf
        \bibitem{expected_performance_bsrt}Fisher, A. Expected Performance of the LHC
            Synchrotron-Light Telescope (BSRT) and Abort-Gap Monitor (BSRA).
            (CERN, 2010, 3), \url{https://cds.cern.ch/record/1261191}

        \bibitem{abort_gap_monitor_thesis}Pacner, P. LHC Abort Gap Monitor
            Acquisition System, Master thesis, (Brno, 2022, 59)

% other sources
% https://indico.cern.ch/event/359810/contributions/853246/attachments/716754/983942/BSRA_BI_TB_Feb2015.pdf
% https://edms.cern.ch/document/2794287/1
%

    \end{thebibliography}
} % end \ifboolexpr