### 2024-05-01 bibliography.tex
Update on Overleaf.

\ifboolexpr{bool{jacowbiblatex}}%
{\printbibliography}{

    \begin{thebibliography}{99}

        % http://cds.cern.ch/record/782076/files/CERN-2004-003-V1.pdf?version=2
        \bibitem{LHC_report} Br\"uning, O. \emph{LHC design report Volume I The LHC
            Main Ring}. Geneva: European Organization for Nuclear Research, c2004-. CERN
            (Series), 2004-003, ISBN 92-9083-224-0.

        \bibitem{beam_dump_system} Filippini, R., Carlier, E., Ducimetière, L., Goddard,
            B. \& Uythoven, J. \emph{Reliability Analysis of the LHC Beam Dumping System}.
            (2005,6), \url{https://cds.cern.ch/record/841092}.

        \bibitem{particles_in_ag}Meddahi, M., Bart Pedersen, S., Boccardi, A.,
            Butterworth, A., Fisher, A., Gianfelice-Wendt, E., Goddard, B., Hemelsoet,
            G., Höfle, W., Jacquet, D., Jaussi, M., Kain, V., Lefevre, T., Shaposhnikova,
            E., Uythoven, J. \& Valuch, D. \emph{LHC Abort Gap Monitoring and Cleaning}.
            (2010,6), \url{https://cds.cern.ch/record/1277630}.

        \bibitem{abort_gap_monitor}De Santis, S., Beche, J., Byrd, J., Placidi, M.,
            Turner, W. \& Zolotorev, M. \emph{Development of an Abort Gap Monitor for the
            Large Hadron Collider}.  (2004), \url{https://cds.cern.ch/record/823343}.

        \bibitem{ttc_distribution} Taylor, B. \emph{TTC distribution for LHC detectors}.
            {\em IEEE Trans. Nucl. Sci.}. \textbf{45} pp. 821-828 (1998),
            \url{https://cds.cern.ch/record/363087}.

        \bibitem{high_sensitivity_measurement} Fischer, C.  \emph{High Sensitivity
            Measurement of the Beam Longitudinal Distribution of the LHC Beams}, 
            (2003,1), LHC-BP-ES-0005, rev. 2.0,
            \url{https://edms.cern.ch/document/328145/2.0}.

        \bibitem{bsra_first_operation}Bart Pedersen, S., Lefevre, T., Bravin, E., Boccardi, A.,
          Goldblatt, A., Jeff, A., Roncarolo, F. \& Fisher, A. \emph{First Operation of
          the Abort Gap Monitors for LHC}.  (2010,5), \url{http://cds.cern.ch/record/1272172}.

        \bibitem{lhcb_asic} Böhner, G., Falvard, A., Lecoq, J., Perret, P. \& Trouilleau,
            C. \emph{Very front-end electronics for the LHCb preshower}. (CERN, 2000, 10), \url{http://cds.cern.ch/record/691512}.

        \bibitem{my_thesis}Pacner, P. \emph{LHC Abort Gap Monitor
            Acquisition System}. Master thesis, (Brno, 2022, 59).

        % \bibitem{Fisher_bsra}Fisher, A. \emph{Expected Performance of the LHC
        %     Synchrotron-Light Telescope (BSRT) and Abort-Gap Monitor (BSRA).}
        %     (CERN,2010), \url{https://cds.cern.ch/record/1261191}

        % % https://cds.cern.ch/record/2295107/files/1638883_171-176.pdf
        % \bibitem{machine_protect}Bravin, E., Belohrad, D., Calvo, E. \& Mazzoni, S. Beam
        %     instrumentation for machine protection.  (2014), \url{https://cds.cern.ch/record/2295107}

% other sources
% https://indico.cern.ch/event/359810/contributions/853246/attachments/716754/983942/BSRA_BI_TB_Feb2015.pdf
% https://edms.cern.ch/document/2794287/1
%

    \end{thebibliography}
} % end \ifboolexpr